Variational Inference for Inverse Reinforcement Learning with Gaussian Processes

### 2019-03-13 mpaper/paper.tex
More writing (and a new plot)

\documentclass{mpaper}
\usepackage{bm}
\usepackage{bbm}
\usepackage{mathtools}
\usepackage{thmtools, thm-restate}
\usepackage[UKenglish]{babel}
\usepackage[UKenglish]{isodate}
\usepackage{tikz}

\usetikzlibrary{shapes,bayesnet,arrows.meta}
\tikzstyle{arrow} = [-{Latex[length=2mm,width=2mm]}]

\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}

@@

Inverse reinforcement learning (IRL)---a problem proposed by Russell in
1998 \cite{DBLP:conf/colt/Russell98}---asks us to find a reward function for a
Markov decision process (MDP) that best explains a set of given demonstrations.
IRL is important because reward functions can be hard to define manually
\cite{DBLP:conf/icml/PieterN04,DBLP:journals/corr/abs-1806-06877}, and rewards
are not entirely specific to a given environment, allowing one to reuse the same
reward structure in previously unseen environments
\cite{DBLP:journals/corr/abs-1806-06877,DBLP:conf/uai/JinDAS17,DBLP:conf/nips/LevinePK11}.
Moreover, IRL has seen a wide array of applications in autonomous vehicle
control \cite{DBLP:journals/ijsr/KimP16,DBLP:journals/ijrr/KretzschmarSSB16} and
learning to predict another agent's behaviour
\cite{DBLP:journals/ai/BogertD18,DBLP:conf/aaai/VogelRGR12,ziebart2008maximum,DBLP:conf/huc/ZiebartMDB08,DBLP:conf/iros/ZiebartRGMPBHDS09}.
Most approaches in the literature (see Section \ref{sec:background}) make a
convenient yet unjustified assumption that the reward function can be expressed
as a linear combination of features. One proven way to abandon this assumption
is by representing the reward function as a Gaussian process (GP)
\cite{DBLP:conf/uai/JinDAS17,DBLP:conf/nips/LevinePK11,DBLP:journals/corr/abs-1208-2112}.
The original approach used maximum likelihood estimation
\cite{DBLP:conf/nips/LevinePK11}, while the goal of this project is to use
variational inference (VI) instead, which learns approximate posterior
probability distributions instead of point estimates. This approach can prove
useful in three major ways:
\begin{enumerate}
\item Modelling full posterior distributions for various parameters can result
  in more precise reward predictions, as the model simply holds more
  information.
\item Having variance estimates for rewards can direct our choice in what data
  should be collected next.
\item An approximate Bayesian treatment of many parameters in the model guards
  against overfitting \cite{DBLP:conf/uai/JinDAS17}.
\end{enumerate}

@@

One way to model rewards without assumptions of linearity is with a
\emph{Gaussian process} (GP). A GP is a collection of random variables, any
finite combination of which has a joint Gaussian distribution
\cite{DBLP:books/lib/RasmussenW06}. We write $r \sim \mathcal{GP}(0,
k)$ to say that $r$ is a GP with mean $0$ and covariance function
$k$. Covariance functions take two state feature vectors as input and quantify
how similar the two states are, in a sense that we would expect them to have
similar rewards.

As training a GP with $n$ data points has a time complexity of
$\mathcal{O}(n^3)$ \cite{DBLP:books/lib/RasmussenW06}, numerous approximation
methods have been suggested, many of which select a subset of data called
\emph{inducing points} and focus most of the training effort on them
\cite{DBLP:journals/corr/abs-1807-01065}. Let $\mathbf{X_u}$ be the matrix of
features at inducing states, $\mathbf{u}$ the rewards at those states. Then the
full joint probability distribution can be factorised as
\begin{equation} \label{eq:full}
  \pfull = p(\mathbf{u}) \times p(\mathbf{r} \mid \mathbf{u}) \times
  p(\mathcal{D} \mid \mathbf{r}),
\end{equation}
where
\begin{align*}
  p(\mathbf{u}) &= \mathcal{N}(\mathbf{u}; \mathbf{0}, \Kuu) \\
                &= \frac{1}{(2\pi)^{m/2}|\Kuu|^{1/2}}\exp \left( -\frac{1}{2} \mathbf{u}^\intercal\Kuu^{-1}\mathbf{u} \right) \\
                &= \exp\left(-\frac{1}{2}\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u} - \frac{1}{2}\log|\Kuu| - \frac{m}{2}\log 2\pi\right)
\end{align*}
is the GP prior \cite{DBLP:books/lib/RasmussenW06}, and $m \in \mathbb{N}$ is
the number of inducing points. The GP posterior is then a multivariate Gaussian
\cite{DBLP:conf/nips/LevinePK11}
\begin{equation} \label{eq:r}
  p(\mathbf{r} \mid \mathbf{u}) =
  \mathcal{N}(\mathbf{r}; \Kru^\intercal\Kuu^{-1}\mathbf{u}, \Krr - \Kru^\intercal\Kuu^{-1}\Kru),
\end{equation}
and $p(\mathcal{D} \mid \mathbf{r})$ is as in Equation \ref{pDr}. The matrices such as
$\Kru$ are called \emph{covariance matrices} and are defined as
$[\Kru]_{i,j} = k(\mathbf{x}_{\mathbf{r},i},
\mathbf{x}_{\mathbf{u},j})$, where $\mathbf{x}_{\mathbf{r},i}$ and
$\mathbf{x}_{\mathbf{u},j}$ denote feature vectors for the $i$th state in
$\mathcal{S}$ and the $j$th state in $\mathbf{X_u}$, respectively
\cite{DBLP:conf/uai/JinDAS17}.

Given this model, data $\mathcal{D}$, and inducing feature matrix
$\mathbf{X_u}$, our goal is then to find optimal values of parameters
$\bm\lambda$, inducing rewards $\mathbf{u}$, and the rewards for all relevant
states $\mathbf{r}$. While the previous paper to consider this IRL model
computed maximum likelihood (ML) estimates for $\bm\lambda$ and $\mathbf{u}$,
and made an assumption that $\mathbf{r}$ in Equation \ref{eq:r} has zero variance
\cite{DBLP:conf/nips/LevinePK11}, we aim to avoid this assumption and use
VI to approximate the full posterior distribution $\posterior$.
\emph{Variational inference} is an approximation technique for probability
densities \cite{blei2017variational}. Let $\approximation$ be our approximating
family of probability distributions for $\posterior$. Then the job of VI is to
optimise the parameters of the approximating distribution in order to minimise
the \emph{Kullback-Leibler} (KL) divergence between the original probability
distribution and our approximation.  KL divergence (asymmetrically) measures how
different the two distributions are, and can be defined as
\cite{blei2017variational}
\begin{align*}
  \DKL{\approximation}{\posterior} &= \mathbb{E}[\log\approximation - \log\posterior ] \\
                                   &= \mathbb{E}[\log\approximation - \log\pfull] \\
                                   &+ \mathbb{E}[\log p(\mathcal{D})].
\end{align*}
The last term is both hard to compute and constant w.r.t. $\approximation$
\cite{blei2017variational}, so we can remove it from our optimisation objective.
The negation of what remains is often called the \emph{evidence lower bound}
(ELBO) and is defined as\footnote{Throughout the proposal, all integrals should
  be interpreted as definite integrals over the entire sample space.}
\cite{DBLP:books/lib/Bishop07,blei2017variational}
\begin{equation} \label{eq:elbo}
  \begin{split}
    \mathcal{L} &= \mathbb{E}\left[ \log \frac{\pfull}{\approximation} \right] \\
    &= \iiint \log \frac{\pfull}{\approximation} \approximation\dx.
  \end{split}
\end{equation}

@@

The project is primarily concerned with investigating how a VI formulation of
the GP IRL model compares against the original ML approach. Most importantly, we
aim to compare how the two algorithms converge both over time and as the number
of demonstrations increases. It would also be interesting to see how close the
approximation of the posterior distribution is to the real thing. Finally, it is
reasonable to conjecture that VI can outperform point estimates when dealing
with more uncertainty, e.g., when experts make mistakes. We can easily
investigate this by adapting how the evaluation data is generated.

@@

\section{Solution}
% TODO: express all lambdas as exponentials (update the derivatives or sth)
% TODO: give an overview of this section. Also, consider renaming it or
% splitting it into multiple sections

@@

In order to properly investigate the difference between VI
and ML estimation, we keep other parts of the model the same.
Namely, we set the covariance function to a version of the automatic relevance
detection kernel \cite{DBLP:conf/nips/LevinePK11,neal2012bayesian}
\begin{align*}
  k(\mathbf{x}_i, \mathbf{x}_j) = \lambda_0\exp\left( \vphantom{\frac{1}{2}} \right. &-\frac{1}{2}(\mathbf{x}_i - \mathbf{x}_j)^\intercal\bm\Lambda(\mathbf{x}_i - \mathbf{x}_j) \\
                                                                                     &- \mathbbm{1}[i \ne j]\sigma^2\tr(\bm\Lambda) \left. \vphantom{\frac{1}{2}} \right),
\end{align*}
where $\lambda_0$ is the overall `scale' factor for how similar or distant the
states are, $\bm\Lambda = \diag(\lambda_1, \dots, \lambda_d)$ is a diagonal
matrix that determines the relevance of each feature (where $d$ denotes the
number of features), $\mathbbm{1}$ is defined as
\[
  \mathbbm{1}[b] = \begin{cases}
    1 & \mbox{if $b$ is true} \\
    0 & \mbox{otherwise},
  \end{cases}
\]
and $\sigma^2$ is set to $10^{-2}/2$ (as the original paper noted that
the value makes little difference to the performance of the algorithm
\cite{DBLP:conf/nips/LevinePK11}). We will write $\bm\lambda = (\lambda_0,
\dots, \lambda_d)^\intercal$ to refer to both $\lambda_0$ and $\bm\Lambda$ at
the same time.

% TODO: write about lambda having to be variational
% TODO: mention the open problem of finding the expected value of the inverse of
% a matrix, where each element is a non-trivial function (w.r.t. a parameter of
% the function)
%TODO: introductory sentence for the following assumption:
\begin{equation} \label{eq:approximation}
  \approximation = q(\mathbf{u}) \times q(\mathbf{r} \mid \mathbf{u}),
\end{equation}
where
\begin{equation}
  q(\mathbf{u}) = \mathcal{N}(\mathbf{u}; \bm\mu, \bm\Sigma),
\end{equation}
where $\bm\mu \in \mathbb{R}^{m}$ is the mean vector and the $m \times m$
positive semi-definite matrix $\bm\Sigma$ is the covariance matrix
\cite{DBLP:conf/nips/ChengB17,DBLP:journals/jmlr/HensmanDS17,DBLP:conf/aaai/HoangHL17}.

@@

The resulting model is summarised in Figure \ref{fig:graphical_model}. We rely
on $p(\mathcal{D} \mid \mathbf{r})$ as the only link between data and the model.
Since the expression for $q(\mathbf{r} \mid \mathbf{u})$ has both $\mathbf{u}$
and covariance matrices in it, $\mathbf{r}$ depends on both $\mathbf{u}$ and the
parameters of the kernel, $\bm\lambda$. The two remaining dependencies are quite
clear, given that the approximating distribution for $\mathbf{u}$ is
$\mathcal{N}(\bm\lambda, \mathbf{BB}^\intercal)$. % TODO: concluding sentence

@@

In order to derive the ELBO, let us go back to Equation \ref{eq:elbo} and
write
\[
  \mathcal{L} = \mathbb{E}[\log\pfull] - \mathbb{E}[\log\approximation].
\]
By substituting in Equations \ref{eq:full} and \ref{eq:approximation}, we get
\begin{align*}
  \mathcal{L} &= \mathbb{E}[\log p(\mathbf{u}) + \log p(\mathbf{r} \mid \mathbf{u}) + \log p(\mathcal{D} \mid \mathbf{r})] \\
              &- \mathbb{E}[\log q(\mathbf{u}) + \log q(\mathbf{r} \mid \mathbf{u})].
\end{align*}
Since $q(\mathbf{r} \mid \mathbf{u}) = p(\mathbf{r} \mid \mathbf{u})$, they
cancel each other out. Also notice that
\begin{align*}
  \mathbb{E}[\log p(\mathbf{u}) - \log q(\mathbf{u})] &= -\DKL{q(\mathbf{u})}{p(\mathbf{u})} \\
                                                      &= -\frac{1}{2} (\tr (\Kuu^{-1}\bm\Sigma) + \bm\mu^\intercal\Kuu^{-1}\bm\mu - m \\
                                                      &+ \log |\Kuu| - \log |\bm\Sigma|),
\end{align*}
by the definition of KL divergence between two multivariate Gaussians
\cite{kl}. Hence,
\begin{align*}
  \mathcal{L} &= \mathbb{E}\left[ \sum_{i=1}^N \sum_{t=1}^T Q_{\mathbf{r}}(s_{i,t}, a_{i,t}) - \V(s_{i,t}) \right] \\
              &- \frac{1}{2} \left(\tr \left( \Kuu^{-1}\bm\Sigma \right) + \bm\mu^\intercal\Kuu^{-1}\bm\mu + \log |\Kuu| - \log |\bm\Sigma| \right).
\end{align*}
Using the expressions for $Q_{\mathbf{r}}$ we get
\begin{align*}
  \mathcal{L} &= \mathbb{E}\left[\sum_{i=1}^N \sum_{t=1}^T r(s_{i,t}) - \V(s_{i,t}) + \gamma\sum_{s' \in \mathcal{S}} \mathcal{T}(s_{i,t}, a_{i,t}, s')\V(s') \right] \\
              &- \frac{1}{2} \left(\tr \left( \Kuu^{-1}\bm\Sigma \right) + \bm\mu^\intercal\Kuu^{-1}\bm\mu + \log |\Kuu| - \log |\bm\Sigma| \right).
\end{align*}
We can simplify $\sum_{i=1}^N\sum_{t=1}^Tr(s_{i,t})$ by defining a new vector
$\mathbf{t} = (t_1, \dots, t_{|\mathcal{S}|})^\intercal$, where $t_i$ is the
number of times the state associated with the reward $r_i$ has been visited
across all demonstrations. Then
\begin{align*}
  \mathbb{E} \left[ \sum_{i=1}^N\sum_{t=1}^Tr(s_{i,t}) \right] &= \mathbb{E}[\mathbf{t}^\intercal\mathbf{r}] = \mathbf{t}^\intercal\mathbb{E}[\mathbf{r}] \\
                                                               &= \mathbf{t}^\intercal\mathbb{E}\left[\Kru^\intercal\Kuu^{-1}\mathbf{u}\right] = \mathbf{t}^\intercal\Kru^\intercal\Kuu^{-1}\bm\mu.
\end{align*}
This allows us to simplify $\mathcal{L}$ to
\begin{align*}
  \mathcal{L} &= \mathbf{t}^\intercal\Kru^\intercal\Kuu^{-1}\bm\mu - \mathbb{E}[v] \\
              &- \frac{1}{2} \left(\tr \left( \Kuu^{-1}\bm\Sigma \right) + \bm\mu^\intercal\Kuu^{-1}\bm\mu + \log |\Kuu| - \log |\bm\Sigma| \right),
\end{align*}
where
\[
  v = \sum_{i=1}^N \sum_{t=1}^T \V(s_{i,t}) - \gamma\sum_{s' \in \mathcal{S}}
  \mathcal{T}(s_{i,t}, a_{i,t}, s')\V(s').
\]

@@

    \draw [arrow] (perturbation) edge (lemma3);
    \draw [arrow] (lebesgue) edge (theorem);
    \draw [arrow] (derivatives) edge (lemma1);
    \draw [arrow] (derivatives) edge (lemma2);
    \draw [arrow] (derivatives) edge (lemma3);
    \draw [arrow] (measurability) edge (theorem);
    \draw [arrow] (boundedness) edge (theorem);
    \draw [arrow] (lemma1) edge (theorem);
    \draw [arrow] (lemma2) edge (theorem);
    \draw [arrow] (lemma3) edge (theorem);
    \draw [arrow] (lemma4) edge (theorem);
  \end{tikzpicture}
  \caption{A graphical representation of dependencies between our theoretical
    results. An arrow from $A$ to $B$ means that $A$ was used to prove $B$.
    Results from the literature are in gray.}
  \label{fig:theorems}
\end{figure}

@@

@@

    Substituting results from Equations \ref{eq:partial_derivatives1} and
    \ref{eq:partial_derivatives2} back into Equation \ref{eq:2_before} gives:

@@

\section{Derivatives}

>>>>

\documentclass{mpaper}
\usepackage{bm}
\usepackage{bbm}
\usepackage{mathtools}
\usepackage{thmtools, thm-restate}
\usepackage[UKenglish]{babel}
\usepackage[UKenglish]{isodate}
\usepackage{tikz}
\usepackage{layouts}

\usetikzlibrary{shapes,bayesnet,arrows.meta,arrows}
%\tikzstyle{myarrow} = [-{Latex[length=2mm,width=2mm]}]

\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{problem}[theorem]{Open Problem}

@@

Inverse reinforcement learning (IRL)---a problem proposed by Russell in
1998 \cite{DBLP:conf/colt/Russell98}---asks us to find a reward function for a
Markov decision process (MDP) that best explains a set of given demonstrations.
IRL is important because reward functions can be hard to define manually
\cite{DBLP:conf/icml/PieterN04,DBLP:journals/corr/abs-1806-06877}, and rewards
are not entirely specific to a given environment, allowing one to reuse the same
reward structure in previously unseen environments
\cite{DBLP:journals/corr/abs-1806-06877,DBLP:conf/uai/JinDAS17,DBLP:conf/nips/LevinePK11}.
Moreover, IRL has seen a wide array of applications in autonomous vehicle
control \cite{DBLP:journals/ijsr/KimP16,DBLP:journals/ijrr/KretzschmarSSB16} and
learning to predict another agent's behaviour
\cite{DBLP:journals/ai/BogertD18,DBLP:conf/aaai/VogelRGR12,ziebart2008maximum,DBLP:conf/huc/ZiebartMDB08,DBLP:conf/iros/ZiebartRGMPBHDS09}.
Most approaches in the literature (see Section \ref{sec:background}) make a
convenient yet unjustified assumption that the reward function can be expressed
as a linear combination of features. One proven way to abandon this assumption
is by representing the reward function as a Gaussian process (GP)
\cite{DBLP:conf/uai/JinDAS17,DBLP:conf/nips/LevinePK11,DBLP:journals/corr/abs-1208-2112}.
The original approach used maximum likelihood estimation
\cite{DBLP:conf/nips/LevinePK11}, whereas we use variational inference (VI)
instead, which learns approximate posterior probability distributions instead of
point estimates. This approach can prove useful in three major ways:
\begin{enumerate}
\item Modelling full posterior distributions for various parameters can result
  in more precise reward predictions, as the model simply holds more
  information.
\item Having variance estimates for rewards can direct our choice in what data
  should be collected next.
\item An approximate Bayesian treatment of many parameters in the model guards
  against overfitting \cite{DBLP:conf/uai/JinDAS17}.
\end{enumerate}

@@

One way to model rewards without assumptions of linearity is with a
\emph{Gaussian process} (GP). A GP is a collection of random variables, any
finite combination of which has a joint Gaussian distribution
\cite{DBLP:books/lib/RasmussenW06}. We write $r \sim \mathcal{GP}(0,
k)$ to say that $r$ is a GP with mean $0$ and covariance function
$k$. \emph{Covariance functions} (also known as \emph{kernels}) take two state
feature vectors as input and quantify how similar the two states are, in a sense
that we would expect them to have similar rewards.

As training a GP with $n$ data points has a time complexity of
$\mathcal{O}(n^3)$ \cite{DBLP:books/lib/RasmussenW06}, numerous approximation
methods have been suggested, many of which select a subset of data called
\emph{inducing points} and focus most of the training effort on them
\cite{DBLP:journals/corr/abs-1807-01065}. Let $\mathbf{X_u}$ be the matrix of
features at inducing states, $\mathbf{u}$ the rewards at those states. Then the
full joint probability distribution can be factorised as
\begin{equation} \label{eq:full}
  \pfull = p(\mathbf{u}) \times p(\mathbf{r} \mid \mathbf{u}) \times
  p(\mathcal{D} \mid \mathbf{r}),
\end{equation}
where
\begin{align*}
  p(\mathbf{u}) &= \mathcal{N}(\mathbf{u}; \mathbf{0}, \Kuu) \\
                &= \frac{1}{(2\pi)^{m/2}|\Kuu|^{1/2}}\exp \left( -\frac{1}{2} \mathbf{u}^\intercal\Kuu^{-1}\mathbf{u} \right) \\
                &= \exp\left(-\frac{1}{2}\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u} - \frac{1}{2}\log|\Kuu| - \frac{m}{2}\log 2\pi\right)
\end{align*}
is the GP prior \cite{DBLP:books/lib/RasmussenW06}, and $m \in \mathbb{N}$ is
the number of inducing points. The GP posterior is then a multivariate Gaussian
\cite{DBLP:conf/nips/LevinePK11}
\begin{equation} \label{eq:r}
  p(\mathbf{r} \mid \mathbf{u}) =
  \mathcal{N}(\mathbf{r}; \Kru^\intercal\Kuu^{-1}\mathbf{u}, \Krr - \Kru^\intercal\Kuu^{-1}\Kru),
\end{equation}
and $p(\mathcal{D} \mid \mathbf{r})$ is as in \eqref{pDr}. The matrices such as
$\Kru$ are called \emph{covariance matrices} and are defined as
$[\Kru]_{i,j} = k(\mathbf{x}_{\mathbf{r},i},
\mathbf{x}_{\mathbf{u},j})$, where $\mathbf{x}_{\mathbf{r},i}$ and
$\mathbf{x}_{\mathbf{u},j}$ denote feature vectors for the $i$th state in
$\mathcal{S}$ and the $j$th state in $\mathbf{X_u}$, respectively
\cite{DBLP:conf/uai/JinDAS17}.

Given this model, data $\mathcal{D}$, and inducing feature matrix
$\mathbf{X_u}$, our goal is then to find optimal values of parameters
$\bm\lambda$, inducing rewards $\mathbf{u}$, and the rewards for all relevant
states $\mathbf{r}$. While the previous paper to consider this IRL model
computed maximum likelihood (ML) estimates for $\bm\lambda$ and $\mathbf{u}$,
and made an assumption that $\mathbf{r}$ in \eqref{eq:r} has zero variance
\cite{DBLP:conf/nips/LevinePK11}, we aim to avoid this assumption and use
VI to approximate the full posterior distribution $\posterior$.
\emph{Variational inference} is an approximation technique for probability
densities \cite{blei2017variational}. Let $\approximation$ be our approximating
family of probability distributions for $\posterior$. Then the job of VI is to
optimise the parameters of the approximating distribution in order to minimise
the \emph{Kullback-Leibler} (KL) divergence between the original probability
distribution and our approximation.  KL divergence (asymmetrically) measures how
different the two distributions are, and can be defined as
\cite{blei2017variational}
\begin{align*}
  \DKL{\approximation}{\posterior} &= \mathbb{E}[\log\approximation - \log\posterior ] \\
                                   &= \mathbb{E}[\log\approximation - \log\pfull] \\
                                   &+ \mathbb{E}[\log p(\mathcal{D})].
\end{align*}
The last term is both hard to compute and constant w.r.t. $\approximation$
\cite{blei2017variational}, so we can remove it from our optimisation objective.
The negation of what remains is often called the \emph{evidence lower bound}
(ELBO) and is defined as\footnote{Throughout the proposal, all integrals should
  be interpreted as definite integrals over the entire sample space.}
\cite{DBLP:books/lib/Bishop07,blei2017variational}
\begin{equation} \label{eq:elbo}
  \begin{split}
    \mathcal{L} &= \mathbb{E}\left[ \log \frac{\pfull}{\approximation} \right] \\
    &= \iiint \log \frac{\pfull}{\approximation} \approximation\dx.
  \end{split}
\end{equation}

@@

@@

\section{Solution}
% TODO: give an overview of this section. Also, consider renaming it or
% splitting it into multiple sections

@@

We keep the covariance function the same as in the work by Levine et al.
\cite{DBLP:conf/nips/LevinePK11}, which is a version of the automatic relevance
detection kernel \cite{DBLP:conf/nips/LevinePK11,neal2012bayesian}:
\begin{align*}
  k(\mathbf{x}_i, \mathbf{x}_j) = \lambda_0\exp\left( \vphantom{\frac{1}{2}} \right. &-\frac{1}{2}(\mathbf{x}_i - \mathbf{x}_j)^\intercal\bm\Lambda(\mathbf{x}_i - \mathbf{x}_j) \\
                                                                                     &- \mathbbm{1}[i \ne j]\sigma^2\tr(\bm\Lambda) \left. \vphantom{\frac{1}{2}} \right).
\end{align*}
Here, $\lambda_0$ is the overall `scale' factor for how similar or distant the
states are, $\bm\Lambda = \diag(\lambda_1, \dots, \lambda_d)$ is a diagonal
matrix that determines the relevance of each feature (where $d$ denotes the
number of features), $\mathbbm{1}$ is defined as
\[
  \mathbbm{1}[b] = \begin{cases}
    1 & \mbox{if $b$ is true} \\
    0 & \mbox{otherwise},
  \end{cases}
\]
and $\sigma^2$ is set to $10^{-2}/2$ (as the original paper noted that
the value makes little difference to the performance of the algorithm
\cite{DBLP:conf/nips/LevinePK11}). We will write $\bm\lambda = (\lambda_0,
\dots, \lambda_d)^\intercal$ to refer to both $\lambda_0$ and $\bm\Lambda$ at
the same time.

Ideally, we would like to model $\bm\lambda$ with an approximating distribution.
However, due to how $p(\mathbf{u})$ has $\Kuu^{-1}$ in its expression, and the
ELBO is defined as an expectation, we are unable to show that the ELBO is
well-defined. More generally, we pose the following problem, which is open to
the best of our knowledge:
\begin{problem}
  Let $\mathbf{A}$ be a $n \times n$ matrix of coefficients, $X$ be a random
  variable, and $\mathbf{M}$ be an $n \times n$ matrix such that $M_{i,j} = f(X,
  A_{i,j})$, where $f$ is an arbitrary function. Under what circumstances does
  $\mathbb{E}[\mathbf{M}^{-1}]$ exist?
\end{problem}
While there are some obvious examples of when the required expected value
exists (e.g., $f(X, A_{i,j}) = A_{i,j}X$ for an invertible $\mathbf{A}$ and many
distributions of $X$), it would be particularly interesting to know whether the
answer is `always'. A proof of such a result would allow us to model
$\bm\lambda$ instead of treating it as a variational parameter, and would thus
guard against overfitting. For now, $\bm\lambda$ will have to be treated as a
variational parameter.

It remains to decide on the model for $\mathbf{u}$ and $\mathbf{r}$. As is
commonly done when applying VI to GPs, we set
\begin{equation} \label{eq:approximation}
  \approximation = q(\mathbf{u}) q(\mathbf{r} \mid \mathbf{u}),
\end{equation}
where $q(\mathbf{r} \mid \mathbf{u}) = p(\mathbf{r} \mid \mathbf{u})$ and
$q(\mathbf{u}) = \mathcal{N}(\mathbf{u}; \bm\mu, \bm\Sigma)$
\cite{DBLP:conf/nips/ChengB17}.

@@

The resulting model is summarised in Figure \ref{fig:graphical_model}. We rely
on $p(\mathcal{D} \mid \mathbf{r})$ as the only link between data and the model.
Since the expression for $q(\mathbf{r} \mid \mathbf{u})$ has both $\mathbf{u}$
and covariance matrices in it, $\mathbf{r}$ depends on both $\mathbf{u}$ and the
parameters of the kernel, $\bm\lambda$. The two remaining dependencies stem from
the fact that the approximating distribution for $\mathbf{u}$ is
$\mathcal{N}(\bm\lambda, \mathbf{BB}^\intercal)$.

As we want to restrict some parameters (namely, $\bm\lambda$ and the diagonal of
$\mathbf{B}$) to positive values, we express them as exponentials and later
adjust their derivatives accordingly. Specifically, we can set $\lambda_i =
e^{\lambda'_i}$ and optimise $\lambda'_i$ using the chain rule:
\[
  \frac{\partial \mathcal{L}}{\partial \lambda'_i} = e^{\lambda'_i}
  \frac{\partial \mathcal{L}}{\partial \lambda_i}.
\]
This way, we restrict $\lambda_i$ to positive values while allowing $\lambda'_i$
to range over $\mathbb{R}$.

Finally, the parameters are initialised as follows:
\begin{alignat*}{3}
  \mu_i &\sim \mathcal{U}(0, 1) \quad &&\text{for } i = 1, \dots, m, \\
  \lambda_0 &\sim \chi^2_5, && \\
  \lambda_i &\sim \chi^2_1 \quad &&\text{for } i = 1, \dots, d, \\
  \diag(\mathbf{B}) &\sim \chi^2_4, && \\
  \text{the rest of } \mathbf{B} &\sim \mathcal{N}(0, 1). &&
\end{alignat*}
The initialisation of $\bm\mu$ mirrors the initialisation of $\mathbf{r}$ in
previous work by Levine et al. \cite{DBLP:conf/nips/LevinePK11}. While they have
constant initial values for $\bm\lambda$, we sample from $\chi^2$ distributions
centred around those values ($5$ for $\lambda_0$ and $1$ for any other
$\lambda_i$). The distributions for initial values of $\mathbf{B}$ are simply
set to provide a reasonable spread of positive values for the diagonal, and both
positive and negative values for all other entries in the matrix.

@@

In order to derive the ELBO, let us go back to \eqref{eq:elbo} and
write
\[
  \mathcal{L} = \mathbb{E}[\log\pfull] - \mathbb{E}[\log\approximation].
\]
By substituting in \eqref{eq:full} and \eqref{eq:approximation}, we get
\begin{align*}
  \mathcal{L} &= \mathbb{E}[\log p(\mathbf{u}) + \log p(\mathbf{r} \mid \mathbf{u}) + \log p(\mathcal{D} \mid \mathbf{r})] \\
              &- \mathbb{E}[\log q(\mathbf{u}) + \log q(\mathbf{r} \mid \mathbf{u})].
\end{align*}
Since $q(\mathbf{r} \mid \mathbf{u}) = p(\mathbf{r} \mid \mathbf{u})$, they
cancel each other out. Also notice that
\begin{align*}
  \mathbb{E}[\log p(\mathbf{u}) - \log q(\mathbf{u})] &= -\DKL{q(\mathbf{u})}{p(\mathbf{u})} \\
                                                      &= -\frac{1}{2} (\tr (\Kuu^{-1}\bm\Sigma) + \bm\mu^\intercal\Kuu^{-1}\bm\mu - m \\
                                                      &+ \log |\Kuu| - \log |\bm\Sigma|),
\end{align*}
by the definition of KL divergence between two multivariate Gaussians
\cite{kl}. Hence,
\begin{align*}
  \mathcal{L} &= \mathbb{E}\left[ \sum_{i=1}^N \sum_{t=1}^T Q_{\mathbf{r}}(s_{i,t}, a_{i,t}) - \V(s_{i,t}) \right] \\
              &- \frac{1}{2} \left(\tr \left( \Kuu^{-1}\bm\Sigma \right) + \bm\mu^\intercal\Kuu^{-1}\bm\mu + \log |\Kuu| - \log |\bm\Sigma| \right).
\end{align*}
Using the expressions for $Q_{\mathbf{r}}$ we get
\begin{align*}
  \mathcal{L} &= \mathbb{E}\left[\sum_{i=1}^N \sum_{t=1}^T r(s_{i,t}) - \V(s_{i,t}) + \gamma\sum_{s' \in \mathcal{S}} \mathcal{T}(s_{i,t}, a_{i,t}, s')\V(s') \right] \\
              &- \frac{1}{2} \left(\tr \left( \Kuu^{-1}\bm\Sigma \right) + \bm\mu^\intercal\Kuu^{-1}\bm\mu + \log |\Kuu| - \log |\bm\Sigma| \right).
\end{align*}
We can simplify $\sum_{i=1}^N\sum_{t=1}^Tr(s_{i,t})$ by defining a new vector
$\mathbf{t} = (t_1, \dots, t_{|\mathcal{S}|})^\intercal$, where $t_i$ is the
number of times the state associated with the reward $r_i$ has been visited
across all demonstrations. Then
\begin{align*}
  \mathbb{E} \left[ \sum_{i=1}^N\sum_{t=1}^Tr(s_{i,t}) \right] &= \mathbb{E}[\mathbf{t}^\intercal\mathbf{r}] = \mathbf{t}^\intercal\mathbb{E}[\mathbf{r}] \\
                                                               &= \mathbf{t}^\intercal\mathbb{E}\left[\Kru^\intercal\Kuu^{-1}\mathbf{u}\right] = \mathbf{t}^\intercal\Kru^\intercal\Kuu^{-1}\bm\mu.
\end{align*}
This allows us to simplify $\mathcal{L}$ to
\begin{align*}
  \mathcal{L} &= \mathbf{t}^\intercal\Kru^\intercal\Kuu^{-1}\bm\mu - \mathbb{E}[v] \\
              &- \frac{1}{2} \left(\tr \left( \Kuu^{-1}\bm\Sigma \right) + \bm\mu^\intercal\Kuu^{-1}\bm\mu + \log |\Kuu| - \log |\bm\Sigma| \right),
\end{align*}
where
\[
  v = \sum_{i=1}^N \sum_{t=1}^T \V(s_{i,t}) - \gamma\sum_{s' \in \mathcal{S}}
  \mathcal{T}(s_{i,t}, a_{i,t}, s')\V(s').
\]

@@

    \draw [->] (perturbation) edge (lemma3);
    \draw [->] (lebesgue) edge (theorem);
    \draw [->] (derivatives) edge (lemma1);
    \draw [->] (derivatives) edge (lemma2);
    \draw [->] (derivatives) edge (lemma3);
    \draw [->] (measurability) edge (theorem);
    \draw [->] (boundedness) edge (theorem);
    \draw [->] (lemma1) edge (theorem);
    \draw [->] (lemma2) edge (theorem);
    \draw [->] (lemma3) edge (theorem);
    \draw [->] (lemma4) edge (theorem);
  \end{tikzpicture}
  \caption{A graphical representation of dependencies between our theoretical
    results. An arrow from $A$ to $B$ means that $A$ was used to prove $B$.
    Results from the literature are in gray.}
  \label{fig:theorems}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{elbo_over_gamma}
  \caption{Changes in $\mathcal{L}$ as a function of $\bm\mu$, where $\mu_1 =
    \mu_2 = \mu_3$, and $\gamma = 0, 0.1, \dots, 0.9$. Each of the ten curves
    plots a different value of $\gamma$, ranging from $0$ in dark blue to $0.9$
    in dark red.}
\end{figure}

@@

    Substituting results from \eqref{eq:partial_derivatives1} and
    \eqref{eq:partial_derivatives2} back into \eqref{eq:2_before} gives:

@@

\section{Derivatives of the ELBO}
